Optimizing Abstract Abstract Machines

### 2012-11-13 paper/paper.tex

Let $wn : \widehat{State} \times Store \to State$ be the metafunction
that adds a store to a storeless state and $c.\msto$ projects the
store out of a state $c$.

\begin{align*}
(cs, \msto) &\mathrel{\widehat{\machstep}} (cs', \msto') \\
\mbox{where } cs' &= \{c' \mid wn(c, \msto) \machstep wn(c, \msto^c), c \in cs\} \\
              \msto' &= \bigsqcup\{\msto^c \mid wn(c,\msto)\machstep wn(c', \msto^c), c \in cs\}
\end{align*}
That is, the next set of states is what we get by stepping each
given state, without its store. The next store is the combination of all produced
stores. For simplicity, assume that stuck states step to themselves.

@@

To achieve this, we instead allocate an address for the value position
when we create the continuation.  This address and the tail address are
both determined by the label of the application point, so the space
becomes linear and the overall complexity drops to cubic. Note that
this is a critical abstraction in languages with n-ary functions,
since otherwise the continuation space becomes ${\mathcal O}(n^n)$!

>>>>

\begin{align*}
(cs, \msto) &\mathrel{\widehat{\machstep}} (cs', \msto') \\
\mbox{where } cs' &= \{c' \mid wn(c, \msto) \machstep wn(c, \msto^c), c \in cs\} \\
              \msto' &= \bigsqcup\{\msto^c \mid wn(c,\msto)\machstep wn(c', \msto^c), c \in cs\} \\
wn(\ev{\mexp, \menv, \mkont}, \msto) &= \ev{\mexp, \menv, \msto, \mkont} \\
wn(\co{\mval, \mkont}, \msto) &= \co{\mval, \msto, \mkont} \\
wn(\ap{\mvalx{u}, \mval, \mkont}, \msto) &= \ap{\mvalx{u}, \mval, \msto, \mkont}
\end{align*}
That is, the next set of states is what we get by stepping each
given state, without its store. The next store is the combination of all produced
stores. For simplicity, assume that stuck states step to themselves.

@@

To achieve this, we instead allocate an address for the value position
when we create the continuation.  This address and the tail address
are both determined by the label of the application point, so the
space becomes linear and the overall complexity drops to cubic. Note
that this is a critical abstraction in languages with n-ary functions,
since otherwise the continuation space becomes ${\mathcal O}(n^n)$!
This amount of distinction between continuations is unnecessary
in practice.
